 

### 2023-02-10 paper2.Rmd
Adding first cut of thesis
Clean up of formatting and typesetting throughout.

---
title: ''
output: pdf_document

---

\addcontentsline{toc}{section}{Conclusion}
\section*{Conclusion}

Thank you for following this tutorial!

I hope you'll find it useful to write a very professional dissertation.

\cite{ngTracingPathChikungunya2010}

>>>>

```{r loading_libraries}

source("C:/Users/danie/Dropbox (Harvard University)/madagascar_mh/code/0_a_packages_and_paremeters_do_file.R")

```

```{r loading_data}

###### MH DATA

load(file = paste(mh_path, 
                  "mh_data_all_facilities.Rdata", sep = "/"))

###### OUTPUTS

load(file = paste(gpspath_spread,"1_matched",
                  "gps_outputs.Rdata", sep = "/"))

load(file = paste(gpspath_spread,"1_matched", 
                  "gps_facilities.Rdata", sep = "/"))

load(file = paste(resultspath,"3_climate", 
                  "climate_facilities.Rdata", sep = "/"))

load(file = paste(resultspath,"3_climate", 
                  "cyclone_data.Rdata", sep = "/"))

# load(file = paste(analysispath,"2_data", 
#                   "data.Rdata", sep = "/"))

load(file = paste(analysispath,"2_data", 
                  "outputs_models.Rdata", sep = "/"))

load(file = paste(analysispath,"2_data", 
                  "model_outputs_preferred_model.Rdata", sep = "/"))

load(file = paste(analysispath,"2_data", 
                  "outputs_lags.Rdata", sep = "/"))

load(file = paste(analysispath,"2_data", 
                  "model_outputs_regional.Rdata", sep = "/"))

load(file = paste(analysispath,"2_data",
                  "outputs_marginal_effects.Rdata", sep = "/"))

load(file = paste(analysispath,"2_data",
                  "outputs_marginal_effects_predictions.Rdata", sep = "/"))

```

```{r customizing_outputs}

a = "All facilities"
b = "Any mental or neurological disorders"
c = "total"
d = "both"
e = "mean"

num_facilities_CSB <- facilities_all_gps %>% filter(clinic_type_clean == "CSB/FSP") %>% nrow()
num_facilities_CHD <- facilities_all_gps %>% filter(clinic_type_clean == "CHD") %>% nrow()
num_facilities_CHRR_CHU <- facilities_all_gps %>% filter(clinic_type_clean == "CHR/CHU") %>% nrow()
num_facilities <- facilities_all_gps %>% nrow()

num_facilities_CSB = format(num_facilities_CSB, big.mark = ",") 
num_facilities_CHD = format(num_facilities_CHD, big.mark = ",") 
num_facilities_CHRR_CHU = format(num_facilities_CHRR_CHU, big.mark = ",") 
num_facilities_pretty = format(num_facilities, big.mark = ",")

n_facilities_who <- gps_outputs$value[gps_outputs$variable == "n_facilities_who"]
n_facilities_rhino <- gps_outputs$value[gps_outputs$variable == "n_facilities_rhino"]
n_gps_facilities_new <- gps_outputs$value[gps_outputs$variable == "n_gps_facilities_new"]
n_gps_facilities <- gps_outputs$value[gps_outputs$variable == "n_gps_facilities"]

validation_mean_dist <- gps_outputs$value[gps_outputs$variable == "validation_mean_dist"]

validation_max_dist <- gps_outputs$value[gps_outputs$variable == "validation_max_dist"]

n_matched_facilities <- facilities_all_gps %>% filter(source == "Exact - Validated sources") %>% nrow()
n_matched_facilities_RHINO <- facilities_all_gps %>% filter(source == "Exact - RHINoVision") %>% nrow()
n_gps_facilities_fuzzy <- facilities_all_gps %>% filter(source == "Exact - All sources") %>% nrow()
n_centroid <- facilities_all_gps %>% filter(source == "Centroid") %>% nrow()
n_gps_facilities_searches <- facilities_all_gps %>% filter(source == "Searches") %>% nrow()

earliest_climate_date <- min(as.yearmon(start_date_climate), 
                             as.yearmon(as.Date(paste0(as.character(date1), '01'), format='%Y%m%d')))
earliest_climate_date <- format(as.yearmon(earliest_climate_date), "%B %Y")

latest_climate_date <- as.yearmon(end_date_climate)
latest_climate_date <- format(as.yearmon(latest_climate_date), "%B %Y")

n_facilities_sm <- data_climate %>% 
  filter(date == "Jan 2010") %>%
  dplyr::select(sm_monthly, facility_ID) %>%
  filter(!is.na(sm_monthly)) %>%
  dplyr::select(facility_ID) %>%
  unique() %>%
  nrow()

n_facilities_sm_missing <- data_climate %>% 
  filter(date == "Jan 2010") %>%
  dplyr::select(sm_monthly, facility_ID) %>%
  filter(is.na(sm_monthly)) %>%
  dplyr::select(facility_ID) %>%
  unique() %>%
  nrow()

```

```{r dynamic_text_results}

function_dynamic_text_results <-
  function(section = "main",         #c(main, lags, regional, multiverse)
           coefficient_option,
           lag_option,
           region_option = "all",
           model_option = "4",
           value = "irr",
           value_include = TRUE,
           multiverse_sign = "positive",
           multiverse_p = FALSE) #(TRUE, FALSE, only),) 
  {
    output_df <- get(paste0(
      "model_outputs",
      ifelse(
        section == "main",
        "_preferred_model",
        ifelse(section == "multiverse", "",

               paste0("_", section)))))

    if(section == "multiverse"){

      if(multiverse_sign == "positive"){

        output_df <- output_df %>% filter(estimate >= 0)  %>% filter(coefficient == coefficient_option)

      } else {output_df <- output_df %>% filter(estimate < 0)  %>% filter(coefficient == coefficient_option)}

      if(multiverse_p != FALSE)
      {
        models_counts <- output_df <- output_df %>% filter(p_value < multiverse_p) %>% nrow()
        return(models_counts)

      } else {

        models_counts <- output_df %>% nrow()

        return(models_counts)

      }

    }else

    {

      if(section == "lags"){

        output_df <-  output_df %>% filter(lag == lag_option) %>% filter(coefficient == coefficient_option)

      }

      if(section == "regional")

      {

        output_df <-  output_df %>% filter(coefficient == coefficient_option)  %>%
          filter(region == region_option)
      }

      if(section != "lags" &  section != "regional"){

        output_df <-
          output_df %>% filter(coefficient == coefficient_option) %>%
          filter(model_number == model_option)

      }

      if (value == "irr"){

        val <- round(exp(output_df$estimate), 3)
        p_val <- round(output_df$p_value, 3)
        low <- round(exp(output_df$lower_99), 3)
        high <- round(exp(output_df$upper_99), 3)

        if (value_include == TRUE){

          return(paste0(
            "(IRR: ",
            val,
            ", p ",
            ifelse(p_val < 0.001,
                   "< 0.0001, ",
                   paste0("= ",
                          p_val,
                          ", ")),
            "99\\% CI: ",
            low,
            " – ",
            high
            ,
            ")"))

        }

        if (value_include == FALSE){

          return(paste0(
            "(p ",
            ifelse(p_val < 0.001,
                   "< 0.0001, ",
                   paste0("= ",
                          p_val,
                          ", ")),
            "99\\% CI: ",
            low,
            " – ",
            high
            ,
            ")"))

        }

        if (value_include == "only"){
          # IRR only

          return(paste0(ifelse(
            val < 1,
            paste0(round(1 - val, 3) * 100, "\\% fewer"),
            paste0(round(val - 1, 3) * 100, "\\%")
          )))

        }

      }

      if (value == "ame"){
        val <- round(output_df$ame, 3)
        p_val <- round(output_df$ame_p_value, 3)
        low <- round(output_df$ame_lower_99, 3)
        high <- round(output_df$ame_upper_99, 3)

        if (value_include == TRUE){
          return(paste0(
            "(AME: ",
            val,
            ", p ",
            ifelse(p_val < 0.001,
                   "< 0.001, ",
                   paste0("= ",
                          p_val,
                          ", ")),
            "99\\% CI: ",
            low,
            " – ",
            high
            ,
            ")"))

        }

        if (value_include == FALSE) {
          # AME without value

          return(paste0(
            "(p ",
            ifelse(p_val < 0.001,
                   "< 0.001, ",
                   paste0("= ",
                          p_val,
                          ", ")),
            "99\\% CI: ",
            low,
            " – ",
            high
            ,
            ")"))

        }

        if (value_include == "only") {
          # AME only

          return(round(val, 2))

        }
      }
    }
  }

num_models_multiverse = format(max(model_outputs$model_number), big.mark = ",")

function_pval_stars <- function(x){

  ifelse(stars.pval(x) ==".", "", stars.pval(x))

  }

```

# Climate exposures and population mental health in Madagascar: an ecological study of national health information system data

\begin{centering}

\emph{Daniel Arias, Karestan Koenen, Christopher Golden}

\vspace{12pt}

\noindent\rule{8cm}{0.4pt}

\end{centering}

\begin{raggedright}

{\bf Background:} Climate change-related exposures (CCEs) have been increasingly linked to worsened mental health, yet few studies of CCEs have taken place in low- and middle-income countries.

{\bf Methods:} To better understand how climate conditions may be impacting population mental health in these settings, we leverage meteorologic, geospatial, and health system data reported by `r num_facilities_pretty` facilities in Madagascar from 2010 to 2020 to explore the mental health impacts of three exposures of climate using negative binomial regression: temperature, soil moisture, and duration of cyclones and tropical storms.

{\bf Findings and interpretation:} We found a statistically significant reduction in monthly cases associated with temperature—approximately `r function_dynamic_text_results(section = "main", coefficient_option = "exposure_temperature", model_option = 4, value = "irr", value_include = "only")` cases reported per each degree Celsius `r function_dynamic_text_results(section = "main", coefficient_option = "exposure_temperature", model_option = 4, value = "irr", value_include = TRUE)`. A multiverse analysis of `r num_models_multiverse` models found this estimate to be robust to alternative specifications, while a subgroup analysis by ecoregion indicated that this association was primarily localized to facilities in the central highlands, where cooler temperatures are more prevalent. We further found that higher soil moisture may be associated with higher case counts in the future, with an `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include = "only")`  elevation in incident cases reported `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include =TRUE)`  after a lag of three months. We further found that greater soil moisture was associated with higher cases reported in the eastern lowlands and the northern dry deciduous forests, which is consistent with the geographic concentration of flood risk in Madagascar.

\vspace*{\fill}

\emph{Keywords: climate mental health, planetary health, eco-anxiety, climate change, humidity, soil moisture, drought, temperature, cyclones.}

\end{raggedright}

\newpage

## Introduction

As empirical evidence increasingly highlights the wide-scale impact of human activities on climate, there is a growing, parallel body of evidence documenting the impacts our changing climate has on human health.[@watts2020ReportLancet2021] Rising temperatures and extreme weather events (including floods, hurricanes, droughts, and fires) have been associated with worsened health outcomes in multiple settings.[@cianconiImpactClimateChange2020a; @rocqueHealthEffectsClimate2021] In particular, these and other climate change-related exposures (CCEs) have been increasingly linked to worsened mental health[@padhyMentalHealthEffects2015  @batterhamSystematicReviewQuantitative2022], with systematic reviews suggesting, among other climate condition impacts, positive associations between high ambient temperatures and mental disorders,[@liuThereAssociationHot2021] drought conditions and suicide,[@stankeHealthEffectsDrought2013] and cyclones and negative mental health outcomes.[@tasdikhasanExploringMentalHealth2020]

Despite the weight of this evidence, few quantitative studies evaluating the effects of climate on population mental health have, to date, taken place in low- and middle-income countries (LMICs),[@sharpeClimateChangeClimaterelated2021] populations that—despite contributing the least to human-driven climate change—are among the most likely to be vulnerable to its impacts.[@bakerProjectingGlobalEmissions2020]

Among low-income countries most vulnerable to the health consequences of climate change is Madagascar, a nation that generates less than a tenth of a percentage of global carbon dioxide emissions. Worsening CCEs in Madagascar (primarily in the form of droughts, floods, and cyclones) are likely to exacerbate endemic, climate-sensitive threats to health, including malnutrition and malaria.[@rakotoarisonAssessmentRiskVulnerability2018] In particular, persistent drought conditions in recent years have led to extensive food insecurity.[@adamtaylorMadagascarHeadedClimate] Though there is debate concerning the extent to which human activity is directly responsible for Madagascar's current drought,[@harringtonAttributionSevereLow; @zhongManyDimensionsDrought2021] the extremity of the situation has led some observers to classify conditions in Madagascar as the first climate change famine in history.[@worldfoodprogrammeMadagascarSevereDrought2021]

Understanding how both incremental and extreme changes in climate impact population mental health is critical to gauging the full extent to which CCEs impact human health and wellbeing under a multidisciplinary approach to population health.[@berryCaseSystemsThinking2018] This understanding is particularly important in contexts such as Madagascar, where climate conditions are already demonstrating pronounced pressures on economic, social, and physical wellbeing.

In this ecological study, we investigated the association of climate conditions and population mental health in Madagascar between 2010 and 2020 using routine health system data from Madagascar’s Ministry of Public Health. The use of health system data enabled us to indirectly evaluate the extent to which existing public health data may already be showing signals of CCEs, which may aid efforts to prevent and mitigate mental health stressors and better design systemic responses to mental health needs. As a measure of population mental health, we considered facility-level monthly counts of new diagnoses of mental disorders. To capture the importance of both variation in temperature and precipitation conditions in Madagascar, we explored the predictive association of three climate conditions with incident mental health disorders: mean monthly ambient air temperature, mean monthly soil moisture, and monthly cyclone activity.

## Methods

### Data

To analyze the impact of climate on population mental health in Madagascar, we first obtained health management information system (HMIS) data from the Madagascar Ministère de la Santé Publique. We then obtained gridded temperature and soil moisture data from the European ReAnalysis (ERA5)[@copernicusclimatechangeservicec3sFifthGenerationECMWF2017] and the European Space Agency’s Climate Change Initiative (ESA CCI),[@europeanspaceagencyEuropeanSpaceAgency2022] respectively. Data on tropical storms and cyclones were obtained from open-source compilations of storms in the South-West Indian Ocean[@SouthWestIndianOcean2023], which were in turn cross-referenced against storm trajectory data from the National Oceanic and Atmospheric Administration (NOAA)[@InternationalBestTrack2021] for accuracy.

#### Outcomes

Madagascar’s health information system provided facility-level data on counts of disease incidence summarized and reported monthly. Prior to 2019, this system was the Gestion du Système d'Information Sanitaire (GESIS); in early 2019, the system transitioned to using District Health Information Software 2 (DHIS2).

Among facilities reporting any mental health data between 2010 and 2020, data were obtained from `r num_facilities_CSB` public and private primary health centers (known as Centres Santé de Base or CSBs and Formations Sanitaires Privées de Base or FSBs, respectively), `r num_facilities_CHD` district referral hospitals (CHDs), `r num_facilities_CHRR_CHU` regional referral hospitals (CHRRs) and university hospitals (CHUs), for a total of `r num_facilities_pretty` uniquely-identified facilities in our sample.

Exploratory data analysis indicated that the classification and reporting of counts of mental disorders in GESIS varied across types of facilities in their level of detail. Among CHD facilities, incidence of mental disorders was aggregated with neurological disorders (coded as “neuro psychiatric disorders”), while CHRR and CHU facilities reported more detailed counts of specific mental disorders, such as depression, personality disorders, and schizophrenia. Across all facilities, disaggregated data was reported by age group but not by sex. Among CSB facilities, incidences of mental disorders were reported by age group under “mental illnesses and psychic disorders” until June 2015, when reporting was reclassified to “mental disorders,” and disaggregated data was reported by both age and sex.

To address heterogeneity in reporting, our analysis aggregated any monthly incidence of mental or neurological disorders reported by a facility, irrespective of age and sex, into a summary outcome measure. To address outliers, monthly observations of the outcome were arranged separately for each facility into a unique time series. The nonparametric Friedman’s super smoother regression estimator[@friedmanVariableSpanScatterplot1984] was applied to each time series to identify outlier observations and replace them through linear interpolation.[@hyndmanAutomaticTimeSeries2008] By considering outliers on a facility-by-facility basis, the data cleaning approach preserved extreme values that may reflect true variances in incidence (for example, incidences reported by a large, regional reference hospital) while addressing potential errors in data entry. (Additional detail on the data cleaning process is available in the Supplementary appendix.)

#### Geospatial data

To retrieve spatial climate data for the facilities in our sample, it was first necessary to geocode facilities to obtain their latitude and longitude. Facility data from GESIS included limited geospatial information. While exact facility coordinates were not available, information on the relevant administrative unit served by a facility was included in the data (e.g., regions for CHRRs, districts for CHDs, and communes for CSBs). These data informed a stepwise matching process to locate facilities in our sample.

First, region, district, and commune names in our sample were standardized against reference shapefiles to address variations in translation, abbreviation, transposition, accent marks, spacing, hyphenation, and spelling. (A summary of the standardized names is provided in the Supplementary appendix.) Next, facility names, types, and associated administrative units were used to match facilities to a validated spatial inventory of health facilities in Sub-Saharan Africa,[@mainaSpatialDatabaseHealth2019] of which `r format(as.numeric(n_facilities_who), big.mark = ",")` coordinates were provided for public facilities in Madagascar. In order to systematically match facilities in our sample to facilities in the spatial inventory, we implemented approximate string matching, which allowed for facility names and administrative units to be matched approximately, rather than exactly, to corresponding patterns in the validated spatial inventory. Applying this matching technique, `r format(as.numeric(n_matched_facilities), big.mark = ",")` facilities (`r round(as.numeric(n_matched_facilities)/num_facilities * 100,0)`%) in our sample were matched to validated coordinates.

Among the remaining facilities, an additional `r format(as.numeric(n_matched_facilities_RHINO), big.mark = ",")` primary facilities were matched to coordinates from a spatial inventory of `r format(as.numeric(n_facilities_rhino), big.mark = ",")` coordinates obtained from the Routine Health Information Network (RHINO),[@rhinovisiondecisionsupportsystemMadagascarDecisionSupport] which supports an open source database of health facility data in conjunction with the United States Agency for International Development, the government of Madagascar, and other stakeholders.

To geolocate the remaining facilities, our existing inventories were combined with additional spatial inventories of health facility coordinates gathered from local consultants, the Global Healthsites Mapping Project, and other sources to create a database of `r format(as.numeric(n_gps_facilities_new), big.mark = ",")` uniquely identified coordinates for potential matches. Using this database, an additional `r format(as.numeric(n_gps_facilities_fuzzy), big.mark = ",")` facilities in our sample were geolocated using manual and approximate string matching.

An additional `r format(as.numeric(n_centroid), big.mark = ",")` facilities were matched to the centroid of their relevant administrative unit in cases where the unit was less than or equal to `r min(commune_lim, fokontany_lim)` square kilometers, as these administrative units are quite small relative to the spatial resolution of gridded climate data (approximately 900 km²). Among the remaining facilities, geolocations were manually identified for `r format(n_gps_facilities_searches, big.mark = ",")` facilities using targeted searches, resulting in `r format(as.numeric(n_matched_facilities) + as.numeric(n_matched_facilities_RHINO) + as.numeric(n_centroid) + as.numeric(n_gps_facilities_fuzzy) + as.numeric(n_gps_facilities_searches), big.mark = ",")` of the `r format(num_facilities, big.mark = ",")` facilities (`r round((as.numeric(n_matched_facilities) + as.numeric(n_matched_facilities_RHINO) + as.numeric(n_centroid) + as.numeric(n_gps_facilities_fuzzy) + as.numeric(n_gps_facilities_searches))/as.numeric(num_facilities) * 100,0)`%) in the sample being geolocated.

To gauge the accuracy of the geolocating process, a subset of 1% of geolocated facilities was selected for validation against manually identified coordinates. Comparing the geolocated coordinates to validated coordinates for each facility, the mean distance between coordinates in the validation sample was `r format(round(as.numeric(validation_mean_dist),2), big.mark = ",")` km., with a maximum distance of approximately `r format(round(as.numeric(validation_max_dist), 1), big.mark = ",")` kilometers, indicating high accuracy in the geolocated process relative to the spatial resolution of climate data. Additional detail on the geolocating and validation process is available in the Supplementary appendix.

#### Climate data

Facility geolocations were then used to retrieve spatially referenced monthly averages of temperature and soil moisture data from ERA5 and ESA CCI, respectively. To allow for modelling of lagged associations, data were obtained for both ERA5 and ESA CCI SM from `r earliest_climate_date` to `r latest_climate_date`.

ERA 5 is a global atmospheric reanalysis from the European Centre for Medium-Range Weather Forecasts (ECMWF) and is freely available through its Copernicus Climate Change Service (C3S)@europeanspaceagencyEuropeanSpaceAgency2022. The spatial resolution of ERA 5 is approximately 30 km., with quality-assured hourly estimates available from 1959 to the present  @europeanspaceagencyEuropeanSpaceAgency2022. Using Google Earth Engine, a time series of monthly aggregate values for mean air temperature in degrees Kelvin at a height of 2 meters was obtained for each facility geolocation  @ERA5MonthlyAggregates2022. Temperature data was then converted to degrees Celsius.

We obtained soil moisture data from the European Space Agency’s soil moisture dataset (ESA CCI SM), the world’s first and most comprehensive, multi-decade, satellite-observed dataset of global soil moisture. The spatial resolution of ESA CCI SM is approximately 26 km., with daily estimates available from 1978 to 2021. Using ESA CCI SM active and passive radiometer data, a time series of monthly aggregate values for mean soil moisture—measured as saturation percentage—was constructed from daily observations at each land grid point within Madagascar. We then applied nearest neighbor matching to link these time series to our geolocated facilities.

For a location to have soil moisture data available in ESA CCI SM, it must fall within one of 244,243 land grid points. For some facilities on Madagascar’s coastline, their location placed them outside of ESA CCI’s spatial coverage, resulting in missing values. To address facilities without any soil moisture data, nearest neighbor searching was conducted to identify the nearest land grid point for which ESA CCI SM data would be available; this nearest neighbor’s soil moisture time series data would then be applied to the facility with missing data. Within a time series of ESA CCI data for a given location, however, monthly data could still be missing if adequate satellite observation was not recorded anytime during a given month; for these values, we used linear interpolation to address missing data within a time series.

In addition, we obtained data on tropical storms and cyclones in the South-West Indian Ocean for the 2009-2010 to 2020-2021 seasons, inclusive. Open source data on tropical cyclones in the south-west Indian Ocean was obtained by scraping information from Wikipedia summaries of each cyclone season, which were originally sourced from alerts issued by Météo-France La Réunion (MFR La Réunion), the World Meteorological Organisation designated Regional Specialized Meteorological Centre (RSMC) for the provision of forecasts and warnings of tropical cyclones in the South-West Indian Ocean. To ensure that the compiled data accurately reflected storms in Madagascar during the study period, we cross-referenced these data against storm trajectory data for the South Indian Basin from NOAA.[@InternationalBestTrack2021]

With these data, we generated two indicators to measure exposure to storm activity in a given month: a binary indicator for whether a storm impacted Madagascar in that month and a count indicator for the number of days per month with a storm. To capture lingering post-dissipation effects, we defined impact as the storm's duration plus an additional 14 days.

### Statistical analysis

All analyses were completed using R (version 4.2.1).[@rcoreteamLanguageEnvironmentStatistical2022] Marginal effects were calculated using the 'margins' package (version 0.3.26)[@leeperMarginsMarginalEffects2021] for R.

#### Modeling approach

Exploratory data analysis highlighted that the variance of monthly facility-level counts of incident mental disorders highly exceeded the mean; consequently, our statistical analysis utilized negative binomial (NB) regression, which is recommended in the case of over-dispersed count data.

Our primary analysis involved conducting four models to estimate the association between facility-level counts of incident mental disorders and climate conditions. Model 1 used temperature as the principal predictor of counts of incident mental disorders. Model 2 only used soil moisture as a predictor. Model 3 used both temperature and soil moisture independently as predictors, while Model 4—our preferred model—introduced an interaction term between the two. 
Duration of cyclone activity was included in all models using our count indicator of exposure. In all models, we included fixed effects for Madagascar’s twenty-two regions to account for geographic differences in climate patterns and case reporting. We further included month and year fixed effects to account for seasonality and time-varying omitted variables.

In addition, to aid in the interpretation of results, our analysis further reported the average marginal effect of each model predictor. Marginal effects clarify the effect of a per-unit change in a continuous explanatory variable (e.g., temperature) on an outcome where the regression model involves interaction. (In models without interaction or higher-order terms, the slope coefficient from the model will equal the marginal effect.) In addition to estimating the average marginal effects of our temperature, soil moisture, and cyclone variables, we further calculated the marginal effect and predicted values of the outcome for each unit of climate exposure across the three variables of interest.

To test for lagged associations—that is, whether in a given month, climate conditions in prior months were associated with counts of mental disorders—versions of Model 4 were conducted using lagged predictors. Measures of association were estimated for each monthly lag increment between 1 month and 18 months, inclusive.

Owing to the large number of observations in our sample, significance was evaluated against an alpha of 0.01, with 99% confidence intervals reported. Coefficients and confidence intervals for our indicators of climate exposure were exponentiated to generated incidence rate ratios (IRRs). Confidence intervals were constructed using heteroskedasticity-robust standard errors.

#### Subgroup analysis

Madagascar is home to a diverse range of terrestrial ecoregions, each of which has unique physical, climatic, and ecological characteristics. These variations in the environment include distinct patterns of rainfall, temperature, and cyclone activity, potentially resulting in effect heterogeneity when examining how these exposures impact human health across different ecoregions.

To explore potential heterogeneity of our effect estimates, we evaluated Models 1 through 4 among facilities in each of Madagascar's four predominant ecoregions: the lowland, humid forests of the east, the subhumid forests of the Central High Plateau, the desert and xeric shrublands of the south, and the northeastern Dry deciduous forests. (Ecoregions are presented in Figure \ref{figure_ecological_zone}.)

```{r figure_ecological_zone, include = TRUE, echo=FALSE, warnings = FALSE,  fig.cap="\\label{figure_ecological_zone}Predominant ecoregions of Madagascar, summarized by administrative region.", fig.width = 8, fig.height = 9, fig.align='center', out.width = '90%'}

ecological_zones <- read_excel(
  paste(data_climate_path,
        "ecological_zones.xlsx",
        sep = "/")) %>%
  rename(ecological_zone = 'ecological_zone (1 = Eastern rainforest; 2 = High Plateau; 3 = Arid/Desert; 4 = Dry Deciduos Forest)',
         REGION_NAM = region_name)

map_mdg_1 %>% inner_join(ecological_zones, by = "REGION_NAM") %>%
  mutate(ecological_zone = ifelse(ecological_zone == 1, "Eastern lowland forests",
                                  ifelse(ecological_zone == 2, "Highland subhumid forests",
                                         ifelse(ecological_zone == 3, "Desert and xeric shrubland", "Dry deciduous forests")))) %>%
  mutate(ecological_zone = factor(ecological_zone, levels = c("Eastern lowland forests",
                                                              "Highland subhumid forests",
                                                              "Desert and xeric shrubland",
                                                              "Dry deciduous forests"))) %>%

  ggplot(aes(geometry = geometry, fill = ecological_zone)) +
  geom_sf(size = 0.2,
          color = col_grey_med,
          alpha = 0.6) +
  geom_sf(data = map_mdg_1_un,
          size = 0.2,
          color = col_black,
          fill = NA) +
  theme_void() +
  theme(text = element_text(family = "Source Sans Pro"),
        legend.position = "right") +
  scale_fill_manual(values = c(col_eco_rainforest,
                               col_eco_highlands,
                               col_eco_arid,
                               col_eco_drydecid),
                    name = "Ecoregion") +
  theme_void()

ggsave(filename = paste0("figure_4b.jpg"),
       plot = last_plot(),
       path = analysis_figures,
       width = 8,
       height = 10,
       bg = "white")

```

#### Sensitivity analysis

In order to evaluate the robustness of our main results to alternative specifications, data processing, and modeling decisions, we generated `r num_models_multiverse` plausible model combinations to conduct a multiverse analysis to identify how sensitive effect estimates and confidence intervals were to different methodological choices.

To generate the model combinations, we first detailed seven methodological choices and their potential specifications. These criteria are outlined in Table 1.

```{r table2_1, echo = FALSE, include = TRUE, out.width = '100%'}

knitr::include_graphics("table2_1.png")

```

\addcontentsline{lot}{table}{Table 2.1: Alternative specifications, data processing, and modeling decisions for multiverse analysis}

First, we considered the role of data cleaning in the outcome data and chose to compare our analysis using the cleaned outcome data against the raw, unadjusted values. Second, we chose to compare our findings using data from all facilities (the full sample) against a sample without district and regional hospital (CSB/FSP facilities only). Primary care facilities have smaller catchment areas than hospitals; consequently, it is more plausible that climate data for primary care facilities will closely reflect climate exposure to patients compared to referral hospitals. Third, we considered the decision to model temperature and soil moisture as an additive or interactive relationship (Model 3 vs. Model 4), with both alternatives included in our multiverse analysis. Fourth, we evaluated our decision to account for overdispersion with a NB model by using a Poisson regression framework as an alternative. Fifth, we evaluated alternative specifications of exposure to cyclone activity, using three options: our binary indicator, our count indicator, and no term for cyclone activity included. Sixth, we toggled whether time fixed effects (month and year) were included. Seventh, we similarly toggled regional fixed effects.

<!-- ```{r table1prep, fig.scap="scap table", fig.cap="Real caption of table", include = TRUE, echo=FALSE, warnings = FALSE} -->

<!-- choice <- c("1. Data cleaning for outcome data", -->
<!--             "1. Data cleaning for outcome data", -->
<!--             "2. Sample", -->
<!--             "2. Sample",  -->
<!--             "3. Relationship of temperature and soil moisture", -->
<!--             "3. Relationship of temperature and soil moisture", -->
<!--             "4. Regression framework", -->
<!--             "4. Regression framework", -->
<!--             "5. Cyclone activity", -->
<!--             "5. Cyclone activity", -->
<!--             "5. Cyclone activity", -->
<!--             "6. Month and year fixed effects", -->
<!--             "6. Month and year fixed effects", -->
<!--             "7. Regional fixed effects", -->
<!--             "7. Regional fixed effects") -->

<!-- criteria <- c("a) Linear interpolation to replace outliers*", -->
<!--               " b) No adjustment (raw data)", -->
<!--               "a) Full sample*", -->
<!--               " b) Primary care facilities only (CSB/FSP)", -->
<!--               "a) Additive model", -->
<!--               "b) Interactive model*", -->
<!--               "a) NB model*", -->
<!--               "b) Poisson model", -->
<!--               "a) Count indicator*", -->
<!--               " b) Binary indicator", -->
<!--               "c) Excluded (i.e., no indicator)", -->
<!--               "a) Included*", -->
<!--               "b) Excluded", -->
<!--               "a) Included*", -->
<!--               "b) Excluded") -->

<!-- table_1 <- cbind(choice, criteria) -->

<!-- table_1 <- as.data.frame(table_1) -->

<!-- table_1 %>% -->
<!--   group_by(choice) %>% -->
<!--   gt(rowname_col = "criteria") %>% -->
<!--   tab_header(title = md("Table 2.1: Alternative specifications, data processing, and modeling decisions for multiverse analysis")) %>% -->
<!--   tab_footnote(footnote = "*: preferred specification. CSB/FSP: Centres Santé de Base and Formations Sanitaires Privées de Base. NB: negative binomial model")  -->

<!-- ``` -->

<!-- ```{r table1, fig.scap="scap table", fig.cap="Real caption of table", include = TRUE, echo=FALSE, warnings = FALSE} -->

<!-- kableExtra::kable((table_1$criteria)) %>% -->
<!--     pack_rows(index = table(fct_inorder(table_1$choice)))  -->

<!-- ``` -->

We then interacted the methodological choices to generate `r num_models_multiverse` plausible models. Using the effect estimates and confidence intervals for these models, we constructed specification curves for monthly temperature, soil moisture, and—for the `r as.numeric(num_models_multiverse)/3` models that included the cyclone count indicator— cyclone activity to visually compare the distribution of plausible estimates under different combinations.

#### Ethics statement
The Institutional Review Board (IRB) of the Harvard T.H. Chan School of Public Health determined that the study was not human subjects research, and that additional review was not required (protocol number: IRB21-1303, determined on October 19, 2021).

## Results
Our geolocated sample of facilities—which is displayed in Figure \ref{figure_gps}—showed geographic diversity and comprehensive coverage across Madagascar, with most geolocations identified through validated sources.

Figure \ref{figure_gps_climate} shows the spatial distribution of temperature and soil moisture across the facilities in our sample during January and June 2015, corresponding to the warmer, wet season and the dryer, cool season, respectively. When evaluated over time, annual climate trends in Madagascar showed relatively stable distribution of average temperatures, while soil moisture had a slight upward shift through 2020 (Figure \ref{figure_climate_distribution}). Data of tropical storm and cyclone activity throughout the study period also exhibited seasonal patterns, with most major storms occurring in the hot season, with a peak between February and May (Figure \ref{figure_cyclone}).

\blandscape

```{r figure_gps, include = TRUE, echo=FALSE, warnings = FALSE,  fig.scap="Map of facilities in Madagascar, by GPS coordinate source and facility type.", out.extra='', fig.cap="\\label{figure_gps}Map of facilities in Madagascar, by GPS coordinate source (left panel) and facility type (right panel).", fig.width = 16, fig.height = 10, fig.align='center', out.height= '90%'}

map_a <- facilities_all_gps %>%
  ggplot(aes(geometry=geometry)) +
  geom_sf(data = map_mdg_3_un,
          size = 0.2,
          color = col_grey_light,
          fill = NA) +
  geom_sf(data = map_mdg_2_un,
          size = 0.2,
          color = col_grey_med,
          fill = NA) +
  geom_sf(data = map_mdg_1_un,
          size = 0.2,
          color = col_black,
          fill = NA) +
  geom_sf(alpha = 0.5, aes(color = source)) +
  theme_void() +
  theme(text = element_text(size = 16, family = "Source Sans Pro"),
        legend.position = "left") +
  scale_color_manual(values = c(col_centroid,
                                col_exact_all,
                                col_exact_rhino,
                                col_exact_val,
                                col_search),
                     name = "GPS type")

map_b <- facilities_all_gps %>%
  mutate(clinic_type_clean = ifelse(clinic_type_clean == "CSB/FSP", "Primary",
                                    ifelse(clinic_type_clean == "CHD", "District hospital",
                                           "Regional/university hospital"))) %>%
  mutate(clinic_type_clean = factor(clinic_type_clean, levels = c("Primary",
                                                                  "District hospital",
                                                                  "Regional/university hospital"))) %>%
  ggplot(aes(geometry=geometry)) +
  geom_sf(data = map_mdg_3_un,
          size = 0.2,
          color = col_grey_light,
          fill = NA) +
  geom_sf(data = map_mdg_2_un,
          size = 0.2,
          color = col_grey_med,
          fill = NA) +
  geom_sf(data = map_mdg_1_un,
          size = 0.2,
          color = col_black,
          fill = NA) +
  geom_sf(alpha = 0.5, aes(color = clinic_type_clean)) +
  theme_void() +
  theme(text = element_text(size = 16, family = "Source Sans Pro"),
        legend.position = "right") +
  scale_color_manual(values = c(col_CSB,
                                col_CHD,
                                col_CHR),
                     name = "Facility type")

combined_plot <- ggarrange(map_a, map_b, ncol = 2)

compass_file <- paste0(analysis_figures, "/compass.png")
# From https://commons.wikimedia.org/wiki/File:Gray_compass_rose.svg

ggdraw() +
  draw_image(
    compass_file, scale = .2, x = .51, y = .02,
    hjust = 1, halign = 1, valign = 0
  )  +
  draw_plot(combined_plot)

ggsave(filename = paste0("figure_1.jpg"),
       plot = last_plot(),
       path = analysis_figures,
       width = 16,
       height = 10)

```

\elandscape
